\section{Scattering Theory}
Scattering processes are the processes that transform an initial state into a final state, by some potential, which we will treat as a time-dependent perturbation. This is how we learn about experimental processes.

\subsection{Scattering as a Time-Dependent Perturbation}
We make the assumption that the Hamiltonian may be written as
\begin{equation}
    H = H_0 + V(\vec{x}),
\end{equation}
with
\begin{equation}
    H_0 = \frac{\vec{p}^2}{2m},
\end{equation}
the kinetic energy operator. The eigenvalues of $H_0$ are 
\begin{equation}
    E_{\vec{k}} = \frac{\hbar^2 \vec{k}^2}{2m}.
\end{equation}
Plane wave eigenvectors of $H_0$, with eigenvalues $E_{\vec{k}}$, are $\k{\vec{k}}$, and we make the assumption that the scattering potential $V(\vec{r})$ is time independent. We will treat this as an incoming particle seeing the scattering potential which is active only when the particle is close to the scatterer. This analysis will be performed using time-dependent perturbation theory in the interaction picture.

We know that the state evolution is
\begin{equation}
    \k{\alpha, t_0; t}_I = U_I(t,t_0)\k{\alpha, t_0;t_0},
\end{equation}
where $U_I(t,t_0)$ is the time evolution operator in the interaction picture, satisfying
\begin{equation}
    i\hbar \partial_t U_I(t,t_0) = V_I(t)U_I(t,t_0).
\end{equation}
We have $U_I(t_0,t_0) = 1$ and $V_I(t)=\exp(iH_0t/\hbar) V \exp(-iH_0t/\hbar)$, and the exact solution to the equation may be written as
\begin{equation}
    U_I(t,t_0) = 1 - \frac{i}{\hbar}\int_{t_0}^t dt' V_I(t')U_I(t',t_0).
\end{equation}
The transition amplitude for the transformation from an initial state $\k{i}$ into a final state $\k{n}$, both eigenstate to $H_0$, can be written as 
\begin{equation}
    \bkop{n}{U_I(t,t_0)}{i} = \delta_{ni} - \frac{i}{\hbar} \sum_m \bkop{n}{V}{m} \int_{t_0}^{t} dt' e^{i \omega_{nm} t'} \bkop{m}{U_I(t',t_0)}{i},
\end{equation}
where $\hbar\omega_{nm} = E_n - E_m$ and $\delta{ni}=\bk{n}{i}$. We arrive here by inserting the completeness relation $\sum_m \kb{m}{m}$, and converting $V_I \to V$. 

In scattering theory we deal with a continuum of states, not discrete ones, we therefore have to adjust the above equation to account for this. We do this by quantizing the scattering states in a cubic volume of side $L$. We have, in the coordinate representation, the wave function
\begin{equation}
    \bk{\vec{x}}{\vec{k}} = \frac{1}{L^{3/2}} e^{i\vec{k}\cdot\vec{x}},
\end{equation}
where we will take $L\to\infty$ at the end. Note that $\bk{\vec{k'}}{\vec{k}} = \delta_{\vec{k'}\vec{k}}$.

The initial and final states exist only asymptomatically, and we therefore have to have $t\to\infty$ and $t_0 \to -\infty$.

We start by doing a first order treatment, where we set $\bkop{m}{U_I(t',t_0)}{i} = \delta_{mi}$, which gives
\begin{equation}
    \bkop{n}{U_I(t,t_0)}{i} = \delta_{ni} - \frac{i}{\hbar} \bkop{n}{V}{i} \int_{t_0}^{t} dt' e^{i\omega_{ni}t'}.
\end{equation}
For the $t\to\infty$ we have Fermi's golden rule, and in order to include the behaviour for $t_0 \to - \infty$ we define a matrix $T_{ni}$ by
\begin{equation}
    \bkop{n}{U_I{t,t_0}}{i} = \delta_{ni} - \frac{i}{\hbar} T_{ni} \int_{t_0}^{t} dt' e^{i\omega_{ni}t' + \epsilon t'}
\end{equation}
with $\epsilon > 0$ and $t \ll 1/\epsilon$, and we need to make sure to take the limit $\epsilon \to 0$ before we take $t \to \infty$. We now define the scattering matrix
\begin{equation}
    S_{ni} = \lim_{t\to \infty} \ab[\lim_{\epsilon\to 0} \bkop{n}{U_I{t,-\infty}}{i}] = \delta_{ni} - \frac{i}{\hbar} T_{ni} \int_{-\infty}^{\infty} dt' e^{i\omega{ni} t'} = \delta_{ni} - 2\pi i \delta(E_n - E_i) T_{ni}.
\end{equation}
We can see that the scattering matrix consists of two parts. One part where the final and initial states are the same, and one part governed by the $T$-matrix, with some sort of scattering.

\subsubsection{Transition Rates and Cross Sections}
We define that transition rate as
\begin{equation}
    w(i\to n) = \partial_t  \abs{\bkop{n}{U_I{t,-\infty}}{i}}^2
\end{equation}
and for $\k{i} \neq \k{n}$, we have
\begin{equation}
    \bkop{n}{U_I(t,-\infty)}{i} = - \frac{i}{\hbar} T_{ni} \int_{-\infty}^{t} dt' e^{i\omega_{ni} t' + \epsilon t'} = -\frac{i}{\hbar} T_{ni} \frac{e^{i\omega_{ni} t + \epsilon t}}{i \omega_{ni} + \epsilon}.
\end{equation}
The transition rate for this is then
\begin{equation}
    w(i \to n) = \partial_t \ab [\frac{1}{\hbar^2} \abs{T_{ni}}^2 \frac{e^{2\epsilon t}}{\omega_{ni}^2 + \epsilon^2} ] = \frac{1}{\hbar^2} \abs{T_{ni}}^2 \frac{2\epsilon e^{2\epsilon t}}{\omega_{ni}^2 + \epsilon^2}.
\end{equation}
Here, we need to take $\epsilon \to 0$ first, before taking $t \to \infty$. If $\omega_{ni}\neq 0$, we see that $w \to 0$, while for $\omega_{ni} = 0$, we get a delta function. We find that 
\begin{equation}
    \frac{e e^{2\epsilon t}}{\omega_{ni}^2 + \epsilon^2} = \pi \delta(\omega_{ni}) = \pi \hbar \delta(E_n - E_i),
\end{equation}
and we get the transmission rate
\begin{equation}
    w(i \to n) = \frac{2\pi}{\hbar} \abs{T_{ni}}^2 \delta(E_n - E_i)
\end{equation}
which is independent of time so the limit $t\to\infty$ is trivial. This is very similar to Fermi's golden rule, but with $T_{ni}$ instead of $V_{ni}$, which is a more general matrix. We can use this transition rate to determine a scattering cross section.

Similarly to Fermi's golden rule, we need to determine the density of final states, $\rho(E_n) = \Delta n /\Delta E_n$, so that we can integrate over the final state energy. We will consider elastic scattering, where $\k{i} = \k{\vec{k}}, \k{n}=\k{\vec{k}'}$, and $\abs{\vec{k}} = \abs{\vec{k}'}= k$. Using the normalization due to our cubic volume, we can write
\begin{equation}
    E_{\vec{n}} = \frac{\hbar^2 \vec{k}'^2}{2 m} = \frac{\hbar^2}{2m} \ab (\frac{2\pi}{L})^2 \abs{\vec{n}}^2,
\end{equation}
and thus
\begin{equation}
    \Delta E_{\vec{n}} = \partial_{\vec{n}} (\Delta E_{\vec{n}}) \Delta\abs{\vec{n}} = \frac{\hbar^2}{m} \ab(\frac{2\pi}{L})^2 \abs{\vec{n}} \Delta \abs{\vec{n}},
\end{equation}
where $\vec{n} = n_x \hv{i} + n_y \hv{j} + n_z \hv{k}$ and $n_{x,y,z} \in \mathbb{Z}$. We have $\vec{n} = (L/2\pi) \abs{\vec{k}'} = (L/2\pi)k$ and $L$ is large, so we can imagine $\abs{\vec{n}}$ as almost continuous. The number of states within a spherical shell of radius $\abs{\vec{n}}$ and thickness $\Delta\abs{\vec{n}}$ is 
\begin{equation}
    \Delta n = 4\pi \abs{\vec{n}}^2 \Delta \abs{\vec{n}} \times \frac{d\Omega}{4\pi},
\end{equation}
considering the fraction of solid angle represented by the final state vector $\vec{k}'$. Thus,
\begin{equation}
    \rho(E_n) = \frac{\Delta n}{\Delta E_n} = \frac{m}{\hbar^2} \ab(\frac{L}{2\pi})^2 \abs{\vec{n}} d\Omega = \frac{mk}{\hbar^2} \ab(\frac{L}{2\pi})^3 d\Omega.
\end{equation}
Integrating over the final states gives
\begin{equation}
    w(i\to n) = \frac{m k L^3}{(2\pi)^2 \hbar^3} \abs{T_{ni}}^2 d\Omega.
\end{equation}

This concept of cross section is used to understand the transition rate in scattering experiments. Determining the rate at which particles scatter into a solid angle $d\Omega$ from a beam of particles with momentum $\hbar\vec{k}$. The speed of the particles is $v = \hbar k /m$, and the time needed to cross the cubic volume is $t = L/v$. The particle flux in the beam is $v / L^3$, and the probability flux for the wave function becomes
\begin{equation}
    \vec{j}(\vec{x}, t) = \ab (\frac{\hbar}{m}) \frac{\vec{k}}{L^3} = \frac{v}{L^3}.
\end{equation}
The cross section $d\sigma$ is defined as the transition rate divided by the flux. Finally, we arrive at
\begin{equation}
    \frac{d\sigma}{d\Omega} = \ab(\frac{mL^3}{2 \pi\hbar^2})^2 \abs{T_{ni}}^2.
\end{equation}
We now have to relate the matrix elements $T_{ni}$ to the scattering potential $V(\vec{r})$.

\subsubsection{Solving for the $T$ matrix}
Recalling the definition of the $T$ matrix we have
\begin{equation}
    \bkop{n}{U_I(t,-\infty)}{i} = \delta_{ni} + \frac{1}{\hbar} T_{ni} \frac{e^{i\omega_{ni}t + \epsilon t}}{-\omega_{ni}+ i\epsilon}
\end{equation}
and 
\begin{equation}
    \bkop{n}{U_I(t,-\infty)}{i} = \delta_{ni} - \frac{i}{\hbar} \sum_m V_{nm} \int_{-\infty}^{t} dt' e^{i\omega_{nm}t'} \bkop{m}{U_I(t', -\infty)}{i}.
\end{equation}
Inserting the former equation into the latter under the integral sign we obtain three terms. The first of which is $\delta_{ni}$, the second looks like the former equation but with $T_{ni}$ replaced by $V_{ni}.$ The final term is
\begin{equation}
    -\frac{i}{\hbar} \frac{1}{\hbar} \sum_m V_{nm} \frac{T_{mi}}{-\omega_{mi} + i\epsilon} \int_{-\infty}^{t} dt' e^{i(\omega_{nm} + \omega_{mi})t' + \epsilon t'}.
\end{equation}
Since $\omega_{nm} + \omega_{mi} = \omega_{ni}$, the result of the integral can be taken out from the summation. Gathering terms and comparing to the former equation we find that
\begin{equation}
    T_{ni} = V_{ni} + \frac{1}{\hbar} \sum_m V_{nm} \frac{T_{mi}}{-\omega_{mi} + i\epsilon} = V_{ni} + \sum_m V_{nm} \frac{T_{mi}}{E_i - E_m + i\hbar\epsilon}.
\end{equation}
This is an inhomogeneous system of linear equation which can be solved for $T_{ni}$ in terms of the matrix elements $V_{nm}$, which are known.

We can define a set of vectors $\k{\psi^{(+)}}$ in terms of components in some basis $\k{j}$, such that
\begin{equation}
    T_{ni} = \sum_j \bkop{n}{V}{j}\bk{j}{\psi^{(+)}} = \bkop{n}{V}{\psi^{(+)} }.
\end{equation}
Inserting this into the equation for $T_{ni}$, we find
\begin{equation}
    \bkop{n}{V}{\psi^{(+)} } = \bkop{n}{V}{i} + \sum_m \bkop{n}{V}{m} \frac{\bkop{m}{V}{\psi^{(+)}}}{E_i - E_m + i\hbar \epsilon}.
\end{equation}
Since this must be true for all $\k{n}$ we find 
\begin{equation}
    \k{\psi^{(+)}} = \k{i} + \frac{1}{E_i - H_0 + i\hbar \epsilon} V\k{\psi^{(+)}},
\end{equation}
which is known as the Lippmann-Schwinger equation. We can define an operator $T$ with matrix elements $T_{ni}$ by writing
\begin{equation}
    T\k{i} = V \k{\psi^{(+)}},
\end{equation}
we can then find the operator equation
\begin{equation}
    T = V + V\frac{1}{E_i - H_0 + i\hbar\epsilon}T. \label{eq:TV}
\end{equation}

\subsubsection{Scattering from the Future to the Past}
We could instead imagine the scattering process evolving backwards in time, from a plane wave state $\k{i}$ in the far future, to a state $\k{n}$ in the distant past. We would in this case write 
\begin{equation}
    U_I(t,t_0) = 1 + \frac{i}{\hbar} \int_{t}^{t_0} dt' V_I(t') U_I(t',t_0)
\end{equation}
and the $T$ matrix is the n defined as 
\begin{equation}
    \bkop{n}{U_I(t,t_0)}{i} = \delta_{ni} + \frac{1}{\hbar} T_{ni} \int_{t}^{t_0} dt' e^{i\omega_{ni} t' - \epsilon t'}.
\end{equation}
In this case we can define $T$ using a different set of states $\k{\psi^{(-)}}$ through
\begin{equation}
    T\k{i} = V \k{\psi^{(-)}}.
\end{equation}

\subsection{The Scattering Amplitude}
We make the substitution $\hbar \epsilon \to \epsilon$ in the Lippmann-Schwinger equation. We continue considering elastic scattering and we use $E$ for the initial (and final) energy. We write
\begin{equation}
    \k{\psi^{(\pm)}} = \k{i} + \frac{1}{E-H_0 \pm i\epsilon} V\k{\psi^{(\pm)}}
\end{equation}
and multiply by $\b{\vec{x}}$ from the left, and by inserting the completeness relation for position states we get
\begin{equation}
    \bk{\vec{x}}{\psi^{(\pm)} } = \bk{\vec{x}}{i} + \int d^3 x' \bkop{\vec{x}}{\frac{1}{E - H_0 \pm i\epsilon}}{\vec{x}'}\bkop{\vec{x}'}{V}{\psi^{(\pm)} },
\end{equation}
which is an integral equation for scattering. The first step is to evaluate the function
\begin{equation}
    G_\pm (\vec{x},\vec{x}') = \frac{\hbar^2}{2m} \bkop{\vec{x}}{\frac{1}{E - H_0 \pm i \epsilon}}{\vec{x}'}.
\end{equation}
The eigenstates of $H_0$ are easier to evaluate in the momentum basis, and we therefore insert complete sets of these. Doing this, we obtain
\begin{equation}
    G_\pm (\vec{x},\vec{x}') = \frac{\hbar^2}{2m} \sum_{\vec{k}'}\sum_{\vec{k}''} \bk{\vec{x}}{\vec{k}'} \bkop{\vec{k}'}{\frac{1}{E-H_0\pm i\epsilon}}{\vec{k}''} \bk{\vec{k}''}{\vec{x}'}
\end{equation}
and letting $H_0$ act on $\k{\vec{k}'}$ we find
\begin{equation}
    \bkop{\vec{k}'}{\frac{1}{E-H_0\pm i\epsilon}}{\vec{k}''} =  \frac{\delta_{\vec{k}'\vec{k}''}}{E - (\hbar^2 \vec{k}'' / 2m) \pm i\epsilon}.
\end{equation}
Recalling
\begin{align}
    \bk{\vec{x}}{\vec{k}'} &= \frac{e^{i\vec{k}'\cdot \vec{x}}}{L^{3/2}}\\
    \bk{\vec{k}''}{\vec{x}'} &= \frac{e^{-i\vec{k}''\cdot \vec{x}'}}{L^{3/2}}
\end{align}
and writing $E = \hbar^2 k^2 / 2m$, we find 
\begin{equation}
    G_\pm (\vec{x}, \vec{x}') = \frac{1}{L^3} \sum_{\vec{k}'} \frac{e^{i\vec{k}'\cdot(\vec{x}-\vec{x}')}}{k^2 - k'^2 \pm i\epsilon}.
\end{equation}
We can convert this to an integral by using
\begin{equation}
    \sum_{\vec{k}} \to \frac{L^3}{(2\pi)^3} \int d^3 k,
\end{equation}
as $L \to \infty$. We now get
\begin{equation}
    G_\pm (\vec{x}, \vec{x}') = \frac{1}{(2\pi)^3} \int d^3 k' \frac{e^{i\vec{k}'\cdot(\vec{x} - \vec{x}')}}{k^2 - k'^2 \pm i \epsilon}.
\end{equation}
This is analytically solvable by switching to spherical coordinates and using contour integration and residuals. Using a contour in either the upper or lower half plane depending on where the singularity is, and applying Cauchy's integral formula, we find
\begin{equation}
    G_\pm (\vec{x}, \vec{x}') = -\frac{1}{4\pi} \frac{e^{\pm i  k\abs{\vec{x} - \vec{x}'}}}{\abs{\vec{x}-\vec{x}'}}
\end{equation}
We recognize that $G_\pm$ is the Green function for the Helmholtz equation
\begin{equation}
    (\nabla^2 + k^2)G_\pm(\vec{x},\vec{x}') = \delta^{(3)}(\vec{x}-\vec{x}'),
\end{equation}
and for $\vec{x}\neq \vec{x}'$ the eigenvalue euqation is solved by $G_\pm$
\begin{equation}
    H_0 G_\pm = E G_\pm.
\end{equation}
Inserting this into the equation for the wave function we find
\begin{equation}
    \bk{\vec{x}}{\psi^{(\pm)}} = \bk{\vec{x}}{i} - \frac{2m}{\hbar^2} \int d^3 x' \frac{e^{\pm ik\abs{\vec{x}-\vec{x}'}}}{4\pi\abs{\vec{x}-\vec{x}'}} \bkop{\vec{x}'}{V}{\psi^{(\pm)}}.
\end{equation}
Note thtat this consists of two terms. The first terms is the incoming wave, while the second is the effect of scattering. Thus, without any scattering, the scattered wave is the same as the incident wave. We can show that the positive solution $\psi^{(+)}$ corresponds to a plane wave plus and outgoing spherical wave, while the negative solution $\psi^{(-)}$ corresponds to a plane wave plus and incoming spherical wave. Which we see as scatterinf forwards or backwards in time. Usually we are interested in the positive solution in physics.

Consider a local potential $V$, that is, a potential only a function of the position operator $\vec{x}$, and diagonal in the position representation. We can write such a local potential as
\begin{equation}
    \bkop{\vec{x}'}{V}{\vec{x}''} = V(\vec{x}') \delta^{(3)}(\vec{x'}-\vec{x}'').
\end{equation}
By inserting a complete set into $\bkop{\vec{x}'}{V}{\psi^{(\pm)}}$ we get
\begin{equation}
    \bkop{\vec{x}'}{V}{\psi^{(\pm)}} = \int d^3 x'' \bkop{\vec{x}'}{V}{\vec{x}''}\bk{\vec{x}''}{\psi^{(\pm)}} = V(\vec{x}')\bk{\vec{x}'}{\psi^{(\pm)}},
\end{equation}
which we can use to simplify the equation for the wave function
\begin{equation}
    \bk{\vec{x}}{\psi^{(\pm)}} = \bk{\vec{x}}{\psi^{(\pm)}} = \bk{\vec{x}}{i} - \frac{2m}{\hbar^2} \int d^3 x' \frac{e^{\pm ik\abs{\vec{x}-\vec{x}'}}}{4\pi\abs{\vec{x}-\vec{x}'}} V(\vec{x}') \bk{\vec{x}'}{\psi^{(\pm)}}. 
\end{equation}
We understand $\vec{x}$ as the vector pointing to the observation point, where we measure, or evaluate, the wave function. The vector $\vec{x}'$ is the vector pointing to the scatterer from the incoming wave. When dealing with a finite range potential, we are interested in studying the system at a point far outside the nonvanishing area of the potential, i.e. we don't want to put a detector inside the scatter potential. Generally this means that we can set
\begin{equation}
    \abs{\vec{x}} \gg \abs{\vec{x}'}.
\end{equation}

We introduce $r=\abs{\vec{x}}$, $r' = \abs{\vec{x}'}$, and $\alpha = \angle (\vec{x}, \vec{x}')$. For $r\gg r'$, we write
\begin{alphaAlign}
    \abs{\vec{x}-\vec{x}'} &= \sqrt{r^2 - 2 r r' \cos\alpha + r'^2}\\
    &= r\sqrt{1 - \frac{2r'}{r} \cos\alpha + \frac{r'^2}{r^2} }\\
    &\approx r - \hv{r} \cdot \vec{x}',
\end{alphaAlign}
where
\begin{equation}
    \hv{r} = \frac{\vec{x}}{\abs{\vec{x}}}, \quad \text{and} \quad \vec{k}' = k \hv{r}.
\end{equation}
With this we write
\begin{equation}
    e^{\pm i k \abs{\vec{x-\vec{x}'}}} \approx e^{\pm i k r} e^{\mp i \vec{k}' \cdot \vec{x}'},
\end{equation}
and for large $r$ it is also valid to replave $1/\abs{\vec{x} - \vec{x}'}$ by $1/r$.

Choosing the initial state as an eigenstate to the free particle Hamiltonian $H_0$, such that $\k{i} = \k{\vec{k}}$, we can write
\begin{alphaAlign}
    \bk{\vec{x}}{\psi^{(+)}} &\overset{r \text{ large}}{\rightarrow} \bk{\vec{x}}{\vec{k}} - \frac{1}{4\pi} \frac{2m}{\hbar^2} \frac{e^{ikr}}{r} \int d^3 x' e^{-i \vec{k}' \cdot \vec{x}'} V(\vec{x}') \bk{\vec{x}'}{\psi^{(+)}}\\
    &= \frac{1}{L^{3/2}} \ab [e^{i\vec{k}\cdot \vec{x}} + \frac{e^{ikr}}{r} f(\vec{k}', \vec{k})].
\end{alphaAlign}
Here we see clearly that the scattered wave includes both the incident plane wave, and the outgoing spherical wave with amplitude $f(\vec{k}', \vec{k})$ given by
\begin{alphaAlign}
    f(\vec{k}',\vec{k}) &= \frac{1}{4\pi} \frac{2m}{\hbar^2} L^3 \int d^3 x' \frac{e^{-i \vec{k}'\cdot\vec{x}'}}{L^{3/2}} V(\vec{x}') \bk{\vec{x}'}{\psi^{(+)}}\\
    &=- \frac{m L^3}{2\pi\hbar^2} \bkop{\vec{k}'}{V}{\psi^{(+)}}.
\end{alphaAlign}
Similarly, we can can show for $\k{\psi^{-}}$ that
\begin{equation}
    \bk{\vec{x}}{\psi^{(-)}} = \frac{1}{L^{3/2}} \ab [ e^{i\vec{k}\cdot \vec{x}} + \frac{e^{-ikr}}{r} f(\vec{k}', \vec{k})]
\end{equation}
and
\begin{equation}
    f(\vec{k}', \vec{k}) = -\frac{m L^3}{2 \pi \hbar^2} \bkop{-\vec{k}'}{V}{\psi^{(-)}}.
\end{equation}
We call $f(\vec{k}', \vec{k})$ the scattering amplitude, and we remark that the differential cross section can be written
\begin{equation}
    \frac{d\sigma}{d\Omega} = \abs{f(\vec{k}',\vec{k})}^2.
\end{equation}

\subsubsection{Wave Packet Description}
This description is not valid for a particle being scattered by a scattering center if we model the particle as a wave packet approaching a scattering center. This is hard, but we notice that if the diemsnions of the wave packet is much larger than the size of the scatterer, i.e. the range of $V$, the plane wave approach is valid.

\subsubsection{The Optical Theorem}
A useful and fundamental relationship in quantum mechanics is the optics theorem, which relates the imaginary part of the forward scattering amplitude $f(\vec{k},\vec{k})$ ($\theta = 0$), to the total cross section $\sigma_\text{tot} = \int d\Omega (d\sigma / d\Omega)$, and it reads
\begin{equation}
    \Im f(\vec{k},\vec{k}) = \frac{k \sigma_\text{tot}}{4\pi}.
\end{equation}
The proof of this theorem start with the Lippmann-Schwinger equation, with $\k{i} \to \k{\vec{k}}$
\begin{equation}
    \k{\psi^{(+)}} = \k{\vec{k}} + \frac{1}{E - H_0 + i \epsilon} V \k{\psi^{(+)}}
\end{equation}
and we can rewrite this to obtain
\begin{equation}
    \k{\vec{k}} =  \k{\psi^{(+)}} - \frac{1}{E - H_0 + i \epsilon} V \k{\psi^{(+)}}.
\end{equation}
We can use this and write
\begin{alphaAlign}
    \bkop{\vec{k}}{V}{\psi^{(+)}} &= \ab [ \b{\psi^{(+)}} - \b{\psi^{(+)}} V \frac{1}{E - H_0 - i\epsilon} ] V \k{\psi^{(+)}}\\
    &= \bkop{\psi^{(+)}}{V}{\psi^{(+)}} - \bkop{\psi^{(+)}}{V \frac{1}{E-H_0-i\epsilon} V}{\psi^{(+)}}.
\end{alphaAlign}
We want to take the imaginary part of both sides of this expression. The first term on the right is the expectation value of a Hermitian operator and is thus real. To calculate the imaginary part of the second term we need to use complex analysis and Cauchy's principle value as we get a singularity when $\epsilon \to 0$. Since $\epsilon > 0$, we use a contour integral with a small semicircle extending down to the lower half plane around the singularity. Let this semicircle have the radius $\delta$, and let the singularity be at $z_0 = x_0 + i\epsilon$. To show the trick we will use, let us consider a complex function $f(z)= x + iy$, and we can then write
\begin{equation}
    \int_{-\infty}^{\infty} dx \frac{f(x)}{x-x_0} = \mathscr{P} \int_{-\infty}^{\infty} dx \frac{f(x)}{x-x_0} + \int_c dz \frac{f(z)}{z-z_0},
\end{equation}
where $c$ denotes the semicircle contour, and the Cauchy principle vlaue is defined as
\begin{equation}
    \mathscr{P} \int_{-\infty}^{\infty} dx \frac{f(x)}{x-x_0} = \lim_{\delta \to 0} \ab\{ \int_{-\infty}^{x_0 - \delta} dx \frac{f(x)}{x-x_0}  + \int_{x_0 + \delta}^{\infty}dx \frac{f(x)}{x-x_0}\}.
\end{equation}
For the second term we can use calculus of residues to evaluate
\begin{equation}
    \int_c dz \frac{f(z)}{z-z_0} = i\pi f(x_0) \quad \text{as }\delta \to 0.
\end{equation}
We now have 
\begin{equation}
    \int_{-\infty}^{\infty} dx \frac{f(x)}{x-x_0} = \mathscr{P} \int_{-\infty}^{\infty} dx \frac{f(x)}{x-x_0} + i\pi f(x_0).
\end{equation}

Returning to the equation at hand, we start by writing
\begin{alphaAlign}
    \lim_{\epsilon \to 0} \ab [\frac{1}{E - H_0 - i\epsilon}] &= \lim_{\epsilon\to 0} \int_{-\infty}^{\infty} dE' \frac{\delta(E-E')}{E' - H_0 - i \epsilon}\\
    &= i\pi \delta(E-H_0),
\end{alphaAlign}
using what we showed earlier, and we have omitted the principle value since it is a real number. Recalling $T\k{\vec{k}} = V\k{\psi^{(+)}}$, we can write
\begin{equation}
    \Im \bkop{\vec{k}}{V}{\psi^{(+)}} = -\pi \bkop{\vec{k}}{T^\dagger \delta(E-H_0)T}{\vec{k}}.
\end{equation}
Taking the imaginary part of the forward scattering amplitude gives
\begin{alphaAlign}
    \Im f(\vec{k},\vec{k}) &= -\frac{mL^3}{2\pi \hbar^2} \Im \bkop{\vec{k}}{V}{\psi^{(+)}}\\
    &= \frac{m L^3}{2 \hbar^2} \bkop{\vec{k}}{T^\dagger \delta(E-H_0)T}{\vec{k}}\\
    &= \frac{m L^3}{2 \hbar^2} \sum_{\vec{k}'} \bkop{\vec{k}}{T^\dagger \delta(E-H_0)}{\vec{k}'} \bkop{\vec{k}'}{T}{\vec{k}}\\
    &= \frac{m L^3}{2\hbar^2} \sum_{\vec{k}'} \abs{\bkop{\vec{k}'}{T}{\vec{k}}}^2 \delta_{E, \frac{\hbar^2 \vec{k}'^2}{2m}}, 
\end{alphaAlign}
where $E=\hbar^2k^2/2m$. To simplify, we start with
\begin{equation}
    \bkop{\vec{k}'}{T}{\vec{k}} = \bkop{\vec{k}'}{V}{\psi^{(+)}} = -\frac{2\pi\hbar^2}{m L^3} f(\vec{k}', \vec{k}),
\end{equation}
and using 
\begin{equation}
    \sum_{\vec{k}} \to \frac{L^3}{(2\pi)^3} \int d^3 k,
\end{equation}
leading to
\begin{alphaAlign}
    \Im f(\vec{k}, \vec{k}) &= \frac{m L^3}{2 \hbar^2} \ab (\frac{2\pi \hbar^2}{m L^3})^2 \sum_{\vec{k}'} \abs{f(\vec{k}', \vec{k})}^2 \delta_{E, \frac{\hbar^2 \vec{k}'^2}{2m}} \\
    &\to \frac{2\pi^2 \hbar^2}{m L^3} \frac{L^3}{(2\pi)^3} \int d^3 k' \abs{f(\vec{k}', \vec{k})}^2 \delta\ab(E -  \frac{\hbar^2 \vec{k}'^2}{2m})\\
    &= \frac{\hbar^2}{4\pi m} \int d^3 k' \frac{d\sigma}{d\Omega_{k'}}\delta\ab(E -  \frac{\hbar^2 \vec{k}'^2}{2m}).
\end{alphaAlign}
We have the volume element $d^3k' = k'^2 dk' d\Omega_{k'}$, so we have
\begin{alphaAlign}
    \Im f(\vec{k},\vec{k}) &= \frac{\hbar^2}{4\pi m} \int d\Omega_{k'} \frac{d\sigma}{d\Omega_{k'}} \int d k' k'^2 \delta\ab(E -  \frac{\hbar^2 \vec{k}'^2}{2m})\\
    &= \frac{\hbar^2}{4\pi m} \frac{m}{k \hbar^2} \int d\sigma \int dk' k'^2 \delta(k' - k)\\
    &= \frac{k}{4\pi} \sigma_\text{tot},
\end{alphaAlign}
where we have used the identity for the delta function
\begin{equation}
    \delta(g(x)) = \frac{\delta(x-x_0)}{\abs{\partial_x g(x_0)}}.
\end{equation}

\subsection{The Born Approximation}
To calculate the scattering amplitude $f(\vec{k}', \vec{k})$ for some given potential $V(\vec{x})$ is done by calculating the matrix elements
\begin{equation}
    \bkop{\vec{k}'}{V}{\psi^{(+)}} = \bkop{\vec{k}'}{T}{\vec{k}}
\end{equation}
which is not easy as we lack the necessary closed analytical expressions.

We will consider the expansion of $T$ as powers of $V$, shown by the operator equation Eq. \eqref{eq:TV}, which reads
\begin{equation}
    T = V + V\frac{1}{E - H_0 - i\epsilon}V + V\frac{1}{E - H_0 - i\epsilon}V\frac{1}{E - H_0 - i\epsilon}V + \cdots.
\end{equation}
If we make the first order Born approximation, i.e. the first term, we get $T = V$, or equivalently $\k{\psi^{(+)}} = \k{\vec{k}}$. This gives
\begin{alphaAlign}
    f^{(1)}(\vec{k}', \vec{k}) &= - \frac{mL^3}{2\pi\hbar^2} \bkop{\vec{k}'}{V}{\psi^{+}}\\
    &= -\frac{mL^3}{2\pi \hbar^2} \int d^3 x' \bk{\vec{k}'}{\vec{x}'} \bkop{\vec{x}'}{V}{\vec{k}}\\
    &= - \frac{m L^3}{2 \pi \hbar^2} \int d^3 x' V(\vec{x}') \bk{\vec{k}'}{\vec{x}'} \bk{\vec{x}'}{\vec{k}}\\
    &= -\frac{m L^3}{2\pi \hbar^2} \int d^3 x' V(\vec{x}') \frac{e^{-i\vec{k}'\cdot\vec{x}'}}{L^{3/2}} \frac{e^{i \vec{k}\cdot \vec{x}'}}{L^{3/2}}\\
    &= - \frac{m}{2\pi \hbar^2} \int d^3x' V(\vec{x}') e^{i(\vec{k}-\vec{k}')\cdot \vec{x}'}.\label{eq:firstOrderBorn}
\end{alphaAlign}
Note that this is, disregarding the prefactor, the three-dimensional Fourier transform of the potential $V$ with respect to $\vec{q} = \vec{k}-\vec{k}'$.

Consider the important case where $V$ is a spherically symmetric potential, implying that $f^{(1)}(\vec{k}', \vec{k})$ is a function of $q = \abs{\vec{q}}$. Since energy conservation implies $\abs{\vec{k}'} = k$, we can write 
\begin{equation}
    q = 2 k \sin\frac{\theta}{2}.
\end{equation}
Using this in the integral we can evaluate by switching to spherical coordinates and letting the $z$-axis align with $\vec{q}$,
\begin{alphaAlign}
    f^{(1)}(\theta) &= -\frac{m}{2\pi \hbar^2} \int d^3 x' V(r) e^{iqr\cos\theta} \\
    &=- \frac{m}{2\pi \hbar^2} \int_0^\infty dr r^2 V(r) \underbrace{\int_0^\pi d\theta' \sin\theta' e^{iqr\cos\theta'}}_{u=\cos\theta', \; du = -\sin\theta d\theta'}  \int_0^{2\pi} d\phi\\
    &=- \frac{m}{\hbar^2} \int_0^\infty dr r^2 V(r) \int_{-1}^{1} du e^{i q r u}\\
    &= - \frac{m}{\hbar^2} \int_0^\infty dr r^2 V(r) \frac{e^{iqr} - e^{-iqr}}{iqr}\\
    &= -\frac{2m}{q\hbar^2} \int_0^\infty dr r V(r) \sin(qr)
\end{alphaAlign}

We now consider scattering by a finite square well 
\begin{equation}
    V(r) = 
        \begin{cases}
            V_0 & r\leq a\\
            0   & r > a.
        \end{cases}
\end{equation}
The integral becomes
\begin{alphaAlign}
    f^{(1)}(\theta) &= - \frac{2m}{q\hbar^2} V_0 \int_0^a dr r\sin(qr)\\
    &= - \frac{2m V_0}{\hbar^2} \frac{1}{q} \ab(  \ab[ r \frac{-\cos(qr)}{q} ]_0^a + \frac{1}{q}\int_0^a dr \cos(qr))\\
    &= - \frac{2m V_0}{\hbar^2} \frac{1}{q} \ab( -a\frac{\cos(qr)}{q} + \frac{\sin(qr)}{q^2})\\
    &= - \frac{2m V_0}{\hbar^2} \frac{a^3}{(qa)^2} \ab(\frac{\sin(qr)}{qa} - \cos(qa)).
\end{alphaAlign}
A nuclear potential is rather well approximated by a square well.

Considering another spherically symmetric potential, the Yukawa potential
\begin{equation}
    V(r) = \frac{V_0 e^{-\mu r}}{\mu r}
\end{equation}
where $V_0$ is independent of $r$ and $1/\mu$ is related to the range of the potential. The scattering amplitude is then
\begin{alphaAlign}
    f^{(1)}(\theta) &= - \frac{2m}{\hbar^2} \frac{V_0}{\mu q} \int_{0}^{\infty} dr r \frac{e^{-\mu r}}{r} \sin(qr)\\
    &= - \frac{2m}{\hbar^2} \frac{V_0}{\mu q} \int_{0}^{\infty} dr e^{-\mu r} \sin(qr)\\
    &= - \frac{2m}{\hbar^2} \frac{V_0}{\mu q} \frac{1}{2i} \int_0^\infty dr \ab( e^{-(\mu - i q)r} - e^{-(\mu + i q)r}  )\\
    &= - \frac{2m}{\hbar^2} \frac{V_0}{\mu q} \frac{1}{2i} \ab( \ab[-\frac{e^{-(\mu - i q)r}}{\mu - i q}]_0^\infty - \ab[-\frac{e^{-(\mu + i q)r}}{\mu + i q}]_0^\infty  )\\
    &=- \frac{2m}{\hbar^2} \frac{V_0}{\mu q} \frac{1}{2i} \ab( \frac{1}{\mu - i q} - \frac{1}{\mu + i q})\\
    &= - \frac{2m}{\hbar^2} \frac{V_0}{\mu q} \frac{q}{\mu^2 + q^2}\\
    &= - \ab(\frac{2 m V_0}{\mu \hbar^2}) \frac{1}{q^2 + \mu^2}.
\end{alphaAlign}
Note that
\begin{equation}
    q^2 = 4 k^2 \sin\frac{\theta}{2} = 2 k^2 (1 - \cos\theta).
\end{equation}

The differential cross section for the Yukawa potential is
\begin{equation}
    \frac{d\sigma}{d\Omega} = \ab(\frac{2 m V_0}{\mu\hbar^2})^2 \frac{1}{\ab[2k^2(1-\cos\theta) + \mu^2]^2}.
\end{equation}
If the ratio $V_0/\mu$ is fixed as we take the limit $\mu \to 0$, we find that we obtain the Coulomb potential. If we have $V_0 / \mu = ZZ'e^2$ as $\mu \to 0$, we find
\begin{equation}
    \frac{d\sigma}{d\Omega} = \frac{(2m)^2 (ZZ'e^2)^2}{\hbar^4} \frac{1}{16 k^4 \sin^4(\theta/2)},
\end{equation}
and if we identify $\abs{\vec{p}} = \hbar k$ and the kinetic energy $E_\text{KE} = \abs{\vec{p}}/2m$, we further find that
\begin{equation}
    \frac{d\sigma}{d\Omega} = \frac{1}{16} \ab(\frac{ZZ'e^2}{E_\text{KE}})^2 \frac{1}{\sin^4(\theta/2)}.
\end{equation}
This equation is exactly the Rutherford scattering cross section which is obtained classically.

In general, if the scattering cross section for a spherically symmetric potential can be approximated with the first Born approximation, we have the following properties:
\begin{enumerate}
    \item $d\sigma/d\Omega$ and $f(\theta)$, is a function of $q$ only, with $f(\theta)$ depending on the energy and scattering angle, only through $2k^2 (1-\cos(\theta))$.
    \item $f(\theta)$ is always real.
    \item $d\sigma/d\Omega$ is independent of the sign of $V$.
    \item For small $k$ ($q$ necessarily small)
            \begin{equation}
                f^{(1)}(\theta) = -\frac{1}{4\pi} \frac{2m}{\hbar^2} \int d^3 x V(r)
            \end{equation}
    involving a volume integral independent of $\theta$. (Unsure about the factor $1/4\pi$)
    \item $f(\theta)$ is small for large $q$ due to rapid oscillation of the integrand.
\end{enumerate}

To investigate the validity of the Born approximation we return to the wave function
\begin{equation}
    \bk{\vec{x}}{\psi^{(+)}} = \bk{\vec{x}}{\vec{k}} - \frac{2m}{\hbar^2} \int d^3 x' \frac{e^{i k \abs{\vec{x} - \vec{x}'}}}{4\pi \abs{\vec{x}- \vec{x}'}} V(\vec{x}') \bk{\vec{x}'}{\psi^{(+)}}.
\end{equation}
The approximation is that $T\approx V$, or equivalently $\k{\psi^{(+)}} = \k{\vec{k}}$. The second term on the right-hand side must be much smaller than the first. Assuming a typical value for the potential to be $V_0$, and it is acting within some range $a$. Writing $r' = \abs{\vec{x} - \vec{x}'}$ and evaluating the integral roughly, we obtain the validity condition
\begin{equation}
    \abs{\frac{2m}{\hbar} (\frac{4\pi}{3}a^3)\frac{e^{ikr'}}{4\pi a} V_0 \frac{e^{i \vec{k}\cdot\vec{x}'}}{L^{3/2}}} \ll \abs{\frac{e^{i\vec{k}\cdot\vec{x}}}{L^{3/2}}}.
\end{equation}
In the low energy limit ($ka \ll 1$), the exponential factors may be replaced by unity, and ignoring factors of order unity, we get the simpler criterion
\begin{equation}
    \frac{m\abs{V_0}a^2}{\hbar^2} \ll 1.
\end{equation}

Consider the Yukawa potential, with the range $a=1/\mu$. The validity condition becomes
\begin{equation}
    \frac{m\abs{V_0}}{\hbar^2 \mu^2} \ll 1.
\end{equation}
Comparing this to the condition for the Yukawa potential to develop a bound state, which can be shown is
\begin{equation}
    \frac{2m\abs{V_0}}{\hbar^2 \mu^2} \geq 2.7,
\end{equation}
where $V_0$ is negative. It is clear that for a potential strong enough to have bound states, we cannot use the first order Born approximation. If we instead consider high energies ($ka \gg 1$), the exponential factors instead oscillate strongly over the integration, and cannot be set to unity. It is possible to show that, the validity condition is
\begin{equation}
    \frac{2m}{\hbar^2} \frac{\abs{V_0}a}{k} \ln(ka) \ll 1.
\end{equation}
For larger $k$, the condition is more easily satisfied. Generally, the Born approximation tends to be better at higher energies.

\subsubsection{The Higher-Order Born Approximation}
Writing $T$ to the second order in $V$, we obtain
\begin{equation}
    T = V + V \frac{1}{E -H_0 + i\epsilon} V.
\end{equation}
We continue the Born approximation with
\begin{equation}
    f(\vec{k}', \vec{k}) \approx f^{(1)}(\vec{k}', \vec{k}) + f^{(2)}(\vec{k}', \vec{k})
\end{equation}
where we have 
\begin{equation}
    f^{(1)}(\vec{k}', \vec{k}) = - \frac{m}{2\pi \hbar^2} \int d^3x' V(\vec{x}') e^{i(\vec{k}-\vec{k}')\cdot \vec{x}'},
\end{equation}
and
\begin{alphaAlign}
    f^{(2)}(\vec{k}',\vec{k}) &= -\frac{1}{4\pi} \frac{2m}{\hbar^2} (2\pi)^3 \int d^3 x' \int d^3 x''\\
    &\quad \bk{\vec{k}'}{\vec{x}'}V(\vec{x}') \bkop{\vec{x}'}{\frac{1}{E- H_0 + i\epsilon}}{\vec{x}''} V(\vec{x}'') \bk{\vec{x}''}{\vec{k}}\\
    &= -\frac{1}{4\pi} \frac{2m}{\hbar^2} \int d^3 x' \int d^3 x'' \\
    &\quad e^{-i \vec{k}'\cdot \vec{x}'} V(\vec{x}') \ab[\frac{2m}{\hbar^2} G_+(\vec{x}', \vec{x}'')]V(\vec{x}'')e^{i \vec{k}\cdot\vec{x}''}.
\end{alphaAlign}
The physical interpretation of the second order is that the incident wave interacts with at $\vec{x}''$ via $V(\vec{x}'')$, then propagates from $\vec{x}''$ to $\vec{x}'$ via $G_+(\vec{x}', \vec{x}'')$, where another interaction occurs via $V(\vec{x}')$. Finally, the wave gets scattered in the direction $\vec{k}'$. We see that the second order is interpreted as a two-step scattering process. We can view the subsequent orders similarly.

\subsection{Phase Shifts and Partial Waves}


\subsubsection{Free-Particle States}
The Hamiltonian for a free particle is only the kinetic energy operator, which commutes with the momentum operator, and $\vec{L}^2$, and $L_z$. We can consider a simultaneous eigenket of $H_0, \vec{L}^2, L_z$, ignoring spin, called the spherical wave state, denoted $\k{E,l,m}$. We can denote a general free-particle state as a superposition of $\k{E,l,m}$ states, in the same way we might regard it as a superposition of $\k{\vec{k}}$, with different $\vec{k}$. Thus, we can analyse a free-particle state, either using a plane wave basis $\{\k{\vec{k}}\}$, or a spherical wave basis $\{\k{E,l,m}\}$.

We want to find the transformation function $\bk{\vec{k}}{E,l,m}$ between these two bases. First consider the normalization for spherical waves
\begin{equation}
    \bk{E',l',m'}{E,l,m} = \delta_{ll'}\delta_{mm'}\delta(E-E').
\end{equation}
We guess the angular dependence to be the same as for position-space wave functions, that is, spherical harmonics, and write
\begin{equation}
    \bk{\vec{k}}{E,l,m} = g_{lE}(k) Y_l^m(\hv{k}),
\end{equation}
where the function $g_{lE}(k)$, will be considered later. To prove this we start by considering a plane wave state propagating in the positive $z$-direction, $\k{k\hv{z}}$. This state has the important property
\begin{equation}
    L_z \k{k\hv{z}} = (xp_y -yp_x)\k{k_x=0, k_y=0, k_z=k} = 0.
\end{equation}
Classically, this is reasonable since $\vec{L}\cdot \vec{p} = (\vec{x}\times\vec{p})\cdot\vec{p} = 0$. We can expand $\k{k\hv{z}}$ as
\begin{equation}
    \k{k\hv{z}} = \sum_{l'} \int dE' \k{E',l',m'=0} \bk{E',l'm'=0}{k\hv{z}}.
\end{equation}
We do not sum over $m'$ since $\bk{E',l',m'}{k\hv{z}}=0$ for $m'\neq 0$. To obtain a general ket $\k{\vec{k}}$, we only have to apply the relevant rotation operator 
\begin{equation}
    \k{\vec{k}} = \mathscr{D}(\alpha =\phi, \beta = \theta, \gamma = 0) \k{k\hv{z}}
\end{equation}
Multiplying by $\b{E,l,m}$ from the left we obtain
\begin{alphaAlign}
    \bk{E,l,m}{\vec{k}} &= \sum_{l'} \int dE' \bkop{E,l,m}{\mathscr{D}(\alpha=\phi, \beta = \theta, \gamma = 0)}{E',l'm'=0}\\
    &\quad \times \bk{E',l',m'=0}{k\hv{z}}\\
    &= \sum_{l'} \int dE' \mathscr{D}_{m0}^{(l')} (\alpha=\phi, \beta=\theta, \gamma=0)  \delta_{ll'} \delta(E-E')\\
    &\quad \times \bk{E',l',m'=0}{k\hv{z}}\\
    &= \mathscr{D}_{m0}^{(l)} (\alpha = \phi, \beta = \theta, \gamma = 0) \bk{E,l,m=0}{k \hv{z}}.
\end{alphaAlign}
Since $\bk{E,l,m=0}{k \hv{z}}$ is independent of the orientation of $\vec{k}$, that it is independent of the angles $\theta, \phi$, we can call it
\begin{equation}
    \bk{E,l,m=0}{k \hv{z}} = \sqrt{\frac{2l+1}{4\pi}} g_{lE}^*(k).
\end{equation}
We recall that we can write a spherical harmonic as
\begin{equation}
    Y_l^{m'*} (\theta, \phi) = \sqrt{\frac{2l+1}{4\pi}} \mathscr{D}_{m'0}^{(l)}(\alpha = \phi, \beta = \theta, \gamma = 0),
\end{equation}
and we can then write
\begin{equation}
    \bk{\vec{k}}{E,l,m} = g_{lE}(k) Y_l^m(\hv{k}).
\end{equation}
We note 
\begin{equation}
    (H_0 - E)\k{E,l,m} = 0,
\end{equation}
and 
\begin{equation}
    \b{\vec{k}} (H_0 - E) = \ab(\frac{\hbar^2 k^2}{2m} - E) \b{\vec{k}}.
\end{equation}
Combining these we see that
\begin{equation}
    \ab(\frac{\hbar^2 k^2}{2m} - E) \bk{\vec{k}}{E,l,m} = 0,
\end{equation}
so $\bk{\vec{k}}{E,l,m}$ can only be non-zero only if $E = \hbar^2 k^2 /2m$, so we are able to write
\begin{equation}
    g_{lE} (k) = N \delta\ab(\frac{\hbar^2 k^2}{2m} - E).
\end{equation} 
Returning to the normalization condition we find
\begin{equation}
    \bk{E',l',m'}{E,l,m} = \abs{N}^2 \frac{mk'}{\hbar^2} \delta(E-E') \delta{ll'}\delta(mm'),
\end{equation}
where we see that $N=\hbar/\sqrt{mk}$ is sufficient. We therefore end up with
\begin{equation}
    \bk{\vec{k}}{E,l,m} = \frac{\hbar}{\sqrt{mk}} \delta\ab(\frac{\hbar^2 k^2}{2m} - E) Y_l^m(\hv{k}).
\end{equation}
From this we can find the expansion of $\k{\vec{k}}$ in the spherical wave states as
\begin{alphaAlign}
    \k{\vec{k}} &= \sum_l \sum_m \int dE \k{E,l,m}\bk{E,l,m}{\vec{k}}\\
    &= \sum_{l=0}^{\infty} \sum_{m=-l}^{l} \k{\frac{\hbar^2 k^2}{2m},l,m} \ab(\frac{\hbar}{\sqrt{mk}} Y_l^{m*}(\hv{k}))  
\end{alphaAlign}

We now consider the wave function in position space, which from wave mechanics is known as
\begin{equation}
    \bk{\vec{x}}{E,l,m} = c_l j_l(kr) Y_l^m(\hv{r}),
\end{equation}
where $j_l(kr)$ is the spherical Bessel function of order $l$. To find $c_l$, we evaluate
\begin{alphaAlign}
    \bk{\vec{x}}{\vec{k}} &= \sum_l \sum_m \int dE \bk{\vec{x}}{E,l,m}\bk{E,l,m}{\vec{k}}\\
    &= \sum_l \sum_m \int dE c_l j_l(kr) Y_l^m(\hv{r}) \frac{\hbar}{\sqrt{mk}} \delta(E - \frac{\hbar^2 k^2}{2m}) Y_l^{m*}(\hv{k})\\
    &= \sum_l \frac{2l+1}{4\pi} P_l(\hv{k}\cdot \hv{r}) \frac{\hbar}{\sqrt{mk}} c_l j_l(kr),
\end{alphaAlign}
where the addition theorem
\begin{equation}
    \sum_m Y_l^m(\hv{r}) Y_l^{m*}(\hv{k}) = \frac{2l+1}{4\pi} P_l(\hv{k}\cdot\hv{r})
\end{equation}
have been used. We can also write
\begin{equation}
    \bk{\vec{x}}{\vec{k}} = \frac{e^{i\vec{k}\cdot\vec{x}}}{(2\pi)^{3/2}} = \frac{1}{(2\pi)^{3/2}} \sum_l (2l+1)i^l j_l(kr) P_l(\hv{k}\cdot \hv{r}),
\end{equation}
which can be shown using the integral representation of the spherical Bessel function
\begin{equation}
    j_l(kr) = \frac{1}{2i^l} \int_{-1}^{1} d(\cos\theta) e^{ikr\cos\theta} P_l(\cos\theta).
\end{equation}
Comparing these two results gives
\begin{equation}
    c_l = \frac{i^l}{\hbar} \sqrt{\frac{2mk}{\pi}}.
\end{equation}
Thus, the wave functions are
\begin{alphaAlign}
    \bk{\vec{k}}{E,l,m} &= \frac{\hbar}{\sqrt{mk}} \delta\ab(\frac{\hbar^2 k^2}{2m} - E) Y_l^m(\hv{k})\\
    \bk{\vec{x}}{E,l,m} &= \frac{i^l}{\hbar} \sqrt{\frac{2mk}{\pi}} j_l(kr) Y_l^m(\hv{r}),
\end{alphaAlign}
which are very useful when developing partial-wave expansion.

\subsubsection{Partial-Wave Expansion}
We return to the case of $V\neq 0$, and assume a spherically symmetric potential. Then, the transition operator $T$ is a scalar operator that commutes with $\vec{L}^2$ and $\vec{L}$.

Using the spherical wave basis, the Wigner-Eckart theorem gives
\begin{equation}
    \bkop{E',l',m'}{T}{E,l,m} = T_l(E) \delta_{ll'} \delta_{mm'},
\end{equation}
where we see that $T$ is diagonal in both $l$ and $m$, and that the diagonal element depends on $E$ and $l$ but not $m$. 

Considering the scattering amplitude
\begin{alphaAlign}
    f(\vec{k}',\vec{k}) &= -\frac{1}{4\pi} \frac{2m}{\hbar^2} L^3 \bkop{\vec{k}'}{T}{\vec{k}}\\
    &\to - \frac{1}{4\pi}\frac{2m}{\hbar} (2\pi)^3 \sum_l\sum_m \sum_{l'}\sum_{m'} \int dE \int dE'\\
    &\quad \times \bk{\vec{k}'}{E',l',m'} \bkop{E',l',m'}{T}{E,l,m} \bk{E,l,m}{\vec{k}}\\
    &= -\frac{1}{4\pi} \frac{2m}{\hbar^2} (2\pi)^3 \frac{\hbar^2}{mk} \sum_l \sum_m T_l\ab(\frac{\hbar^2 k^2}{2m}) Y_L^m(\hv{k}')Y_l^{m*} (\hv{k})\\
    &= -\frac{4\pi^2}{k} \sum_l \sum_m T_l\ab(\frac{\hbar^2 k^2}{2m}) Y_l^m(\hv{k}') Y_l^{m*}(\hv{k}). 
\end{alphaAlign}
Choosing a coordinate system such that $\vec{k}$ is in the positive $z$-direction, we have
\begin{equation}
    Y_l^m(\hv{k}) = \sqrt{\frac{2l+1}{4\pi}} \delta_{m0}.
\end{equation}
Taking $\theta$ to be the angle between $\vec{k}$ and $\vec{k}'$, we can write
\begin{equation}
    Y_l^0(\vec{k}') = \sqrt{\frac{2l+1}{4\pi}} P_l(\cos\theta).
\end{equation}
We define a partial-wave amplitude $f_l(k)$ as
\begin{equation}
    f_l(k) = -\frac{\pi T_l(E)}{k}.
\end{equation}
We can now rewrite the scattering amplitude to
\begin{equation}
    f(\vec{k}', \vec{k}) = f(\theta) = \sum_{l=0}^{\infty}(2l+1) f_l(k)P_l(\cos\theta).
\end{equation}

Studying the large-distance behaviour of the wave function $\bk{\vec{x}}{\psi^{(+)}}$, using what is found above, gives
\begin{equation}
    \bk{\vec{x}}{\psi^{(+)}} = \frac{1}{(2\pi)^{3/2}} \sum_l (2l+1) \frac{P_l}{2ik} \ab[ (1 + 2ik f_l(k))\frac{e^{ikr}}{r} - \frac{e^{-i(kr-l\pi)}}{r} ].
\end{equation}
We can now interpret the physics of scattering. When a scatterer is absent, we can model the plane wave as a sum of spherically incoming and outgoing waves. The presence of a scatter changes the coefficient of the outgoing waves by
\begin{equation}
    1 \to 1 + 2ik f_l(k),
\end{equation}
leaving the incoming wave unaffected.

\subsubsection{Unitarity and Phase Shifts}
We will now investigate the consequences of unitarity, or equivalently probability conservation. For a time independent formulation the flux current density $\vec{j}$ must satisfy
\begin{equation}
    \nabla \cdot \vec{j} = -\partial_t \abs{\psi}^2 = 0.
\end{equation}
Consider a spherical surface of large radius, then by Gauss' theorem we have
\begin{equation}
    \int_S \vec{j} \cdot d\vec{S} = 0.
\end{equation}
Physically, these two equations tells us that there are no sources or sinks of particles, and the outgoing and incoming flux must be the same.
Due to angular-momentum conservation this must hold for each partial wave separately, or in other words, we must have the coefficient of $e^{ikr}/r $ have the same magnitude as the coefficient of $e^{-ikr}/r$. We define $S_l(k)$ as
\begin{equation}
    S_l(k) = 1 + 2 i k f_l(k)
\end{equation}
and this is then
\begin{equation}
    \abs{S_l(k)} = 1.
\end{equation}
Thus, the only thing that can happen is a change of phase, not magnitude, of the outgoing wave. This equation is known as the unitarity relation for the $l$th partial wave. We can also regard $S_l(k)$ as the $l$th diagonal element of the $S$ operator, which has unitarity as a requirement due to probability conservation.

We can therefore at large distances, at most see a phase change of the outgoing wave in the wave function as a result of scattering. By convention, we call this phase $2\delta_l$, and we write
\begin{equation}
    S_l = e^{2 i \delta_l},
\end{equation}
where $\delta$ is real. We note that $\delta_l$ is an implicit function of $k$. We can now rewrite the partial-wave amplitude as
\begin{equation}
    f_l = \frac{S_l -1}{2ik},
\end{equation}
or in terms of $\delta_l$
\begin{equation}
    f_l = \frac{e^{2i \delta_l} - 1}{2ik} = \frac{e^{i\delta_l}\sin\delta_l}{k} = \frac{1}{k\cot\delta_l - ik},
\end{equation}
depending on what is convenient. The full scattering amplitude is
\begin{alphaAlign}
    f(\theta) &= \sum_{l=0} (2l+1) \ab(\frac{e^{2i\delta_l}-1}{2ik}) P_l(\cos\theta)\\
    &= \frac{1}{k} \sum_{l=0} (2l+1) e^{i\delta_l} \sin\delta_l P_l(\cos\theta).
\end{alphaAlign}
This expression rests on the principles of rotational invariance and probability conservation. It is possible to obtain the same expression by solving the Schrödinger equation with a real, spherically symmetric potential. However, our treatment is of interest as it can more easily be generalized to situations where nonrelativistic potentials fails.

We find the total cross section by
\begin{alphaAlign}
    \sigma_\text{tot} &= \int d\Omega \frac{d\sigma}{d\Omega} = \int d\Omega \abs{f(\theta)}^2\\
    &= \frac{1}{k^2} \int_0^{2\pi} d\phi \int_{-1}^{1} d(\cos\theta) \sum_l \sum_{l'}(2l+1)(2l'+1) e^{i\delta_{l}}\sin\delta_l e^{-i\delta_{l'}}\sin\delta_{l'} P_l P_{l'}\\
    &= \frac{4\pi}{k^2} \sum_l (2l+1)\sin^2\delta_l.
\end{alphaAlign}

As $\delta_l$ is a function of energy, so is $f_l(k)$. The partial wave amplitude, has the restriction of the unitarity relation, which we can explore by drawing an Argand diagram for $kf_l$, noting that
\begin{equation}
    k f_l = \frac{i}{2} + \frac{1}{2} e^{-(i\pi/2) + 2 i \delta_l}.
\end{equation}
We find a circle with radius $1/2$ centred at $z=i/2$, called the unitary circle, on which $kf_l$ must lie. If $\delta_l$ is small we find
\begin{equation}
    f_l = \frac{e^{i\delta_l}\sin\delta_l}{k} \approx \frac{(1+i\delta_l)\delta_l}{k} \approx \frac{\delta_l}{k},
\end{equation}
and $f_l$ is thus almost purely real. If we instead consider $\delta_l$ close to $\pi/2$ we have that $f_l$ is almost purely imaginary with maximized magnitude, and the $l$th partial wave can be considered to be in resonance. The maximum partial cross section is
\begin{equation}
    \sigma_\text{max}^{(l)} = \frac{4\pi}{k^2} (2l+1),
\end{equation}
and is achieved when $\sin^2\delta_l = 1$.

\stepcounter{subsubsection}\subsubsection{Hard-Sphere Scattering}


\stepcounter{subsection}\subsection{Low-Energy Scattering and Bound States}

\subsubsection{Rectangular Well or Barrier}