\section{Symmetry in Quantum Mechanics}

\subsection{Symmetries, Conservation Laws, and Degeneracies}
\subsubsection{Symmetries in Classical Physics}
We will review the concept of symmetry and conservation in classical physics. Suppose we have a Lagrangian $L$, a function of the generalized coordinate $q_i$ and the corresponding velocity $\dot{q}_i$. For a displacement
\begin{equation}
    q_i \to q_i + \delta q_i,
\end{equation}
if the Lagrangian remains unchanged, we have
\begin{equation}
    \frac{\partial L}{\partial q_i} = 0.
\end{equation}
From the Lagrange equation it then follows that 
\begin{equation}
    \frac{d p_i}{dt} = 0,
\end{equation}
where the canonical momentum is defined by
\begin{equation}
    p_i = \frac{\partial L}{\partial \dot{q}_i}.
\end{equation}
So the canonical momentum conjugate to $q_i$ is a conserved quantity under displacement if $L$ remains unchanged.

Similarly, looking at the Hamiltonian formulation we have $H$ as a function of $q_i$ and $p_i$, we have
\begin{equation}
    \frac{d p_i}{dt} = 0
\end{equation}
when 
\begin{equation}
    \frac{\partial H}{\partial t} = 0.
\end{equation}
Thus, if we lack an explicit $q_i$ dependence in $H$, ($H$ has a symmetry under $q_i \to q_i + \delta q_i$) we have a conserved quantity.

\subsubsection{Symmetry in Quantum Mechanics}
In quantum mechanics, we have made the association between an operation, like translation or rotation, to a unitary operator, like $\mathscr{S}$. Regardless if the physical system has a symmetry corresponding to the operator $\mathscr{S}$, we tend to call this operator for a symmetry operator. Furthermore, we have seen that if the symmetry operation is infinitesimal we can write
\begin{equation}
    \mathscr{S} = 1 - \frac{i \epsilon}{\hbar} G,
\end{equation}
where $G$ is the Hermitian generator of the symmetry operator. 

Consider that $H$ is invariant under $\mathscr{S}$. From this, it follows that
\begin{equation}
    \mathscr{S}^\dagger H \mathscr{S} = H,
\end{equation}
or equivalently
\begin{equation}
    \comm{G}{H} = 0.
\end{equation}
From Heisenberg's equation of motion ($i\hbar \partial_t A = \comm{A}{H}$),  we obtain
\begin{equation}
    \frac{d G}{dt} = 0,
\end{equation}
and thus, $G$ is a constant of the motion. E.g. if the Hamiltonian is translationally invariant, the momentum (that generates the translation), is a constant of the motion. 

Consider the system in an eigenket of $G$, when $G$ commutes with $H$, at time $t_0$. For a later time $t>t_0$, we have the ket
\begin{equation}
    \k{g',t_0;t} = U(t,t_0) \k{g'},
\end{equation}
which is also an eigenket of $G$ the same eigenvalue $g'$. Thus, when a ket is an eigenket to $G$ with eigenvalue $g'$ it will stay as such. Proving this is just realizing that since $\comm{G}{H} = 0$, then $\comm{G}{U} = 0$.

\subsubsection{Degeneracies}
Suppose that for a symmetry operator $\mathscr{S}$ we have
\begin{equation}
    \comm{H}{\mathscr{S}} = 0,
\end{equation}
and $\k{n}$ is an energy eigenket with eigenvalue $E_n$. Evidently, $\mathscr{S}\k{n}$ is also an energy eigenket with the same energy. If $\k{n}$ and $\mathscr{S}\k{n}$ represent different states we say that the states are degenerate.  Often, the symmetry operator is a function of a continuous parameter, say $\lambda$, and thus all states of the form $\mathscr{S}(\lambda)\k{n}$, are degenerate.

Consider that the symmetry operator in question is rotation, that is we have $\mathscr{D}(R)$. We also suppose that the Hamiltonian is rotationally invariant
\begin{equation}
    \comm{\mathscr{D}(R)}{H} = 0.
\end{equation}
This implies the commutation relations
\begin{alphaAlign}
    \comm{\vec{J}}{H} &= 0\\
    \comm{\vec{J}^2}{H} &= 0,
\end{alphaAlign}
and we can have simultaneous eigenkets of $H$, $\vec{J}^2$, and $J_z$, denoted by $\k{n; j, m}$. All states on the form
\begin{equation}
    \mathscr{R} \k{n; j, m}
\end{equation}
are degenerate and share the same energy. During rotation, different $m$ values are mixed, such that generally, $\mathscr{D}(R)\k{n; j, m}$ is a linear combination of $2j +1$ independent states. We can write this as
\begin{equation}
    \mathscr{D}(R)\k{n; j, m} = \sum_{m'} \k{n; j, m'} \mathscr{D}^{(j)}_{m'm}(R).
\end{equation}
This degeneracy is equal to the number of $m$-values, that is $2j+1$. It is also possible to reach this conclusion by applying $J_\pm$ to the ket.

As an example of this consider an atomic electron subject to the potential $V = V(r) + V_{LS}(r) \vec{L} \cdot \vec{S}$. Since this potential is rotationally invariant, we expect a $(2j+1)$-fold degeneracy. If we applied an external magnetic field, this symmetry would break, and we would get distinct energy levels. 

\subsubsection{SO(4) Symmetry in the Coulomb Potential}
An archetypal example of continuous symmetry in quantum mechanics is the solution of the coulomb potential in the hydrogen atom. The energy eigenvalue for this potential is
\begin{equation}
    E_n = \frac{1}{2} mc^2 \frac{Z^2 \alpha^2}{n^2} = \qty{-13.6}{\electronvolt} \frac{Z^2}{n^2},
\end{equation}
and since this energy is only dependent on $n$, the degeneracy for the ket $\k{n,l,m}$ is 
\begin{equation}
    \sum_{l=0}^{n-1}(2l+1) = n^2.
\end{equation}
This is due to the symmetry of $1/r$ potentials.

Classically, the orbits of $1/r$ potentials have been studied, namely the Kepler problem. The result from this leads us to the conclusion of the existence of a constant of the motion, which in classical mechanics is
\begin{equation}
    \vec{M} = \frac{\vec{p}\times\vec{L}}{m} - \frac{Ze^2}{r}\vec{x},
\end{equation}
called the Lenz vector, or the Runge-Lenz vector.

Treating this quantum mechanically, this new symmetry is called SO(4), analogous to the SO(3) group, but instead rotation operators in four spatial dimensions. We start by modifying the Lenz vector such that it becomes Hermitian
\begin{equation}
    \vec{M} = \frac{1}{2m} (\vec{p}\times\vec{L} - \vec{L}\times\vec{p}) - \frac{Z e^2}{r}\vec{x}.
\end{equation}
We then have
\begin{equation}
    \comm{\vec{M}}{H} = 0,
\end{equation}
for the Hamiltonian
\begin{equation}
    H = \frac{\vec{p}^2}{2m} - \frac{Ze^2}{r},
\end{equation}
and thus $\vec{M}$ is a constant of the motion. Furthermore, the relations
\begin{align}
    \vec{L}\cdot\vec{M} &= \vec{M}\cdot\vec{L} = 0,\\
    \vec{M}^2 &= \frac{2}{m} H (\vec{L}^2 + \hbar^2) + Z^2 e^4, \label{eq:M2}
\end{align}
are useful.

We want to identify the responsible symmetry for this constant of the motion. This is done by exploring the algebra of the generators of this symmetry. We know
\begin{equation}
    \comm{L_i}{L_j} = i\hbar \epsilon_{ijk} L_k,
\end{equation}
and it is possible to show
\begin{equation}
    \comm{M_i}{L_j} = i\hbar \epsilon_{ijk} M_k.
\end{equation}
With this we can derive that
\begin{equation}
    \comm{M_i}{M_j} = -i\hbar \epsilon_{ijk} \frac{2}{m} H L_k.
\end{equation}
These commutation relations do not form a closed algebra as a consequence of the presence of $H$, making it challenging identifying these operators as the generators. Conversely, we will consider the problem with specific bound states. The vector space will now only contain the eigenstates to $H$ that have an energy eigenvalue of $E < 0$, and we may replace $H$ with $E$, closing the algebra. We introduce a scaled vector operator 
\begin{equation}
    \vec{N} = \sqrt{-\frac{m}{2E}} \vec{M},
\end{equation}
such that the closed algebra reads as
\begin{alphaAlign}
    \comm{L_i}{L_j} &= i\hbar \epsilon_{ijk} L_k\\
    \comm{N_i}{L_j} &= i\hbar \epsilon_{ijk} N_k\\
    \comm{N_i}{N_j} &= i\hbar \epsilon_{ijk} L_k,
\end{alphaAlign}

The question becomes what type of symmetry is generated by the operators $\vec{L}$ and $\vec{N}$? Interestingly, the answer is rotation in four spatial dimensions. The first clue for this is that the required number of operators for rotation in $n$ dimensions is given by $n(n-1)/2$, and thus rotation in four dimensions require six operators, which we have with $\vec{L}$ and $\vec{n}$. It is less obvious to see that the above algebra is appropriate for this.

We generalize the procedure performed for three dimensions with $\vec{L}$.  We start by rewriting $(x,y,z)$ and $(p_x, p_y, p_z)$ to $(x_1,x_2, x_3)$ and $(p_1, p_2, p_3)$. Consequently, we write
\begin{alphaAlign}
    L_1 = \tilde{L}_{23} &= x_2 p_3 - x_3 p_2,\\
    L_2 = \tilde{L}_{31} &= x_3 p_1 - x_1 p_3,\\
    L_3 = \tilde{L}_{12} &= x_1 p_2 - x_2 p_1.
\end{alphaAlign}
We know invent a fourth spatial dimension $x_4$ and define its conjugate momentum $p_4$ according to the canonical commutation relation, and write
\begin{alphaAlign}
    N_1 = \tilde{L}_{14} &= x_1 p_4 - x_4 p_1,\\
    N_2 = \tilde{L}_{24} &= x_2 p_4 - x_4 p_2,\\
    N_3 = \tilde{L}_{34} &= x_3 p_4 - x_4 p_3.
\end{alphaAlign}
These operators obey the closed algebra defined earlier. That is, the closed algebra we found, is the algebra for rotation in four spatial dimensions.

We continue by investigating the degeneracies in the Coulomb potential. Starting by defining the operators
\begin{align}
    \vec{I} &= \frac{\vec{L} + \vec{N}}{2},\\
    \vec{K} &= \frac{\vec{L} - \vec{N}}{2},
\end{align}
we can prove the algebra
\begin{alphaAlign}
    \comm{I_i}{I_j} &= i\hbar \epsilon_{ijk} I_k\\
    \comm{K_i}{K_j} &= i\hbar \epsilon_{ijk} K_k\\
    \comm{I_i}{K_j} &= 0.
\end{alphaAlign}
As such, these operators obey independent angular momentum algebra, and we also have that $\comm{\vec{I}}{I} = \comm{\vec{K}}{H} = 0$. We denote the eigenvalues of $\vec{I}^2$ and $\vec{K}^2$ by $i(i+1)\hbar^2$ and $k(k+1)\hbar^2$ respectively, with $i,k$ taking on positive half-integer values.

We find that 
\begin{alphaAlign}
    \vec{I}^2 - \vec{K}^2 &= \frac{1}{4} \ab[ (\vec{L} + \vec{N})^2 - (\vec{L} - \vec{N})^2 ]\\
    &= \frac{1}{4} \ab[ (\vec{L}^2 + \vec{N}^2 + 2\vec{L}\cdot\vec{N}) - (\vec{L}^2 + \vec{N}^2 - 2\vec{L}\cdot\vec{N}) ]\\
    &= \vec{L} \cdot \vec{N}\\
    &= 0,
\end{alphaAlign}
and thus we must have that $i=k$. Similarly, we find
\begin{alphaAlign}
    \vec{I}^2 + \vec{K}^2 &= \frac{1}{4} \ab[ (\vec{L} + \vec{N})^2 + (\vec{L} - \vec{N})^2 ]\\
    &= \frac{1}{4} \ab[ (\vec{L}^2 + \vec{N}^2 + 2\vec{L}\cdot\vec{N}) + (\vec{L}^2 + \vec{N}^2 - 2\vec{L}\cdot\vec{N}) ]\\
    &= \frac{1}{2} \ab [ \vec{L}^2 + \vec{N}^2 ]\\
    &= \frac{1}{2} \ab [ \vec{L}^2 - \frac{m}{2E} \vec{M}^2 ],
\end{alphaAlign}
that together with Eq. \eqref{eq:M2} gives
\begin{equation}
    2k(k+1)\hbar^2 = \frac{1}{2} \ab ( -\hbar^2 - \frac{m}{2E}  Z^2 e^4).
\end{equation}
Solving this for $E$ yields the relation
\begin{equation}
    E = -\frac{mZ^2 e^4}{2\hbar^2} \frac{1}{(2k+1)^2}.
\end{equation}
This is the energy for the hydrogen atom but with $n$ replaced with $2k+1$. Thus, the rotational symmetries that arise from $\vec{I}$ and $\vec{K}$ creates the degeneracy seen. Technically the degeneracy is $(2i+1)(2k+1) = (2k+1)^2 = n^2$. 

Note here that we found the eigenvalues of the hydrogen atom without solving the Schrödinger equation. Instead, we used the symmetries in the system to obtain the same answer.

The SO(4) algebra used here can also, as can be seen from the above calculations be thought of as SU(2)$\times$SU(2), as $\vec{I}$ and $\vec{K}$ independently follow the SU(2) group algebra.

\subsection{Discrete Symmetries, Parity, or Space Inversion}
Until now, we have considered symmetry operators that are continuous, that is they can be obtained by applying infinitesimal operations successively. We will now consider three types of discrete symmetry operations: Parity, lattice translation, and time reversal.

Consider parity (space inversion). The parity operation, when applied to a coordinate system changes it from a right-handed system to a left-handed system. We will however, consider the parity operator applied to the state kets. Take a general state ket $\k{\alpha}$ and apply the parity operator $\pi$ to obtain a space inverted state
\begin{equation}
    \k{\alpha} \to \pi\k{\alpha}.
\end{equation}
We require from the parity operator that the expectation value of $\vec{x}$ must have the opposite sign when taken with a space-inverted state:
\begin{equation}
    \bkop{\alpha}{\pi^\dagger \vec{x} \pi}{\alpha} = - \bkop{\alpha}{\vec{x}}{\alpha},
\end{equation}
which is to say that we have
\begin{equation}
    \pi^\dagger \vec{x} \pi = - \vec{x},
\end{equation}
or
\begin{equation}
    \vec{x}\pi = -\pi \vec{x}.
\end{equation}
We know that $\pi$ is unitary, and we can see that $\pi$ and $\vec{x}$ must anticommute.

Noting that for an eigenket to the position operator, we have
\begin{equation}
    \vec{x}\pi \k{\vec{x}'} = - \pi \vec{x}\k{\vec{x}'} = (-\vec{x}')\pi \k{\vec{x}'}.
\end{equation}
Thus,  $\pi\k{\vec{x}'}$ is still an eigenket of $\vec{x}$, but with the eigenvalue $-\vec{x}'$, up to a global phase factor, which we generally take to be 1, that is $e^{i\delta} = 1$ for $\delta$ real. With this we see that
\begin{equation}
    \pi^2\k{\vec{x}'} = \k{\vec{x}'},
\end{equation}
and thus $\pi^2 = 1$, and $\pi$ is therefore Hermitian:
\begin{equation}
    \pi^{-1} = \pi^\dagger = \pi,
\end{equation}
with eigenvalue either $+1$ or $-1$.

Consider now the momentum operator $\vec{p}$ as the generator of translation. A translation followed by parity, is equivalent to parity followed by a translation in the reverse direction. Thus,
\begin{equation}
    \pi \mathscr{J}(d\vec{x}') = \mathscr{J}(-d\vec{x}') \pi,
\end{equation}
or equivalently
\begin{equation}
    \pi \ab(1 - \frac{i \vec{p}\cdot d\vec{x}'}{\hbar})\pi^\dagger = 1 + \frac{i\vec{p}\cdot d \vec{x}'}{\hbar}.
\end{equation}
From the above we conclude
\begin{equation}
    \acomm{\pi}{\vec{p}} = 0
\end{equation}
and
\begin{equation}
    \pi^\dagger\vec{p}\pi = -\vec{p}.
\end{equation}

Now take the angular momentum operator $\vec{J}$ as a generator of rotation. For $3\times 3$ orthogonal matrices, we have 
\begin{equation}
    R^{(\text{parity})} R^{(\text{rotation})} =  R^{(\text{rotation})} R^{(\text{parity})},
\end{equation}
for the explicit parity matrix
\begin{equation}
    R^{(\text{parity})} = 
    \begin{pmatrix}
        -1  & 0     & 0\\
        0   & -1    & 0\\
        0   & 0     & -1
    \end{pmatrix}.
\end{equation}
Thus, parity and rotation operations commute. We then postulate the corresponding relation for operators
\begin{equation}
    \pi \mathscr{D}(R) = \mathscr{D}(R)\pi.
\end{equation}
We can now see that 
\begin{equation}
    \comm{\pi}{\vec{J}} = 0
\end{equation}
and
\begin{equation}
    \pi^\dagger \vec{J} \pi = \vec{J}.
\end{equation}
This holds for $\vec{L}$ and $\vec{S}$ individually as well.

Under rotations, $\vec{x}$ and $\vec{J}$ transform as vectors, or spherical tensors of rank 1. However, $\vec{x}$ and $\vec{p}$ are both odd under parity while $\vec{L}$ is even under parity. Odd parity vectors are called polar vectors, while even parity vectors are called axial vectors, or pseudovectors.

Suppose we have an operation on the form $\vec{S}\cdot \vec{x}$, that is a dot product between a polar and axial vector. Under rotation, they behave as a normal scalar (like $\vec{S}\cdot\vec{L}$ or $\vec{x}\cdot\vec{p}$), under parity however, they behave like
\begin{equation}
    \pi^\dagger \vec{S}\cdot\vec{x} = - \vec{S}\cdot \vec{x},
\end{equation} 
while for normal scalars we have
\begin{equation}
    \pi^\dagger \vec{L}\cdot\vec{S}\pi = \vec{L}\cdot\vec{S}.
\end{equation}
We call the operator $\vec{S}\cdot\vec{x}$ for a pseudoscalar.

\subsubsection{Wave Functions and Parity}
We consider a spinless particle with the wave function
\begin{equation}
    \psi(\vec{x}') = \bk{\vec{x}'}{\alpha}.
\end{equation}
For a space inverted state, that is a state with the ket $\pi\k{\alpha}$, has the wave function
\begin{equation}
    \bkop{\vec{x}'}{\pi}{\alpha} = \bk{-\vec{x}'}{\alpha} = \psi(-\vec{x}').
\end{equation}
If we take $\k{\alpha}$ to be an eigenket of parity, we have
\begin{equation}
    \pi\k{\alpha} = \pm\k{\alpha}.
\end{equation}
The wave function is then
\begin{equation}
    \bkop{\vec{x}'}{\pi}{\alpha} = \pm \bk{\vec{x}'}{\alpha}.
\end{equation}
This combined with the earlier wave function gives, depending on the parity of $\k{\alpha}$,
\begin{equation}
    \psi(-\vec{x}') = \pm \psi(\vec{x}') 
        \begin{cases}
            + & \text{even parity}\\
            - & \text{odd parity}.
        \end{cases}
\end{equation}

Note that not all wave functions that are of interest for physical systems follow a parity law like the one above. For example, since the momentum operator anticommutes with the parity operator, the momentum eigenket is not a parity eigenket. We can see that the plane wave function does not obey the parity relation above.

Eigenkets of the orbital angular momentum however, are expected to be parity eigenkets since $\comm{\pi}{\vec{L}} = 0$. We investigate this by considering the wave function under space inversion
\begin{equation}
    \bk{\vec{x}'}{\alpha; l, m} = R_\alpha(r) Y_l^m(\theta, \phi).
\end{equation}
The parity transformation $\vec{x}' \to -\vec{x}'$ is accomplished for spherical coordinates by
\begin{alphaAlign}
    r &\to r\\
    \theta &\to \pi - \theta \quad (\cos\theta \to -\cos\theta)\\
    \phi &\to \phi + \pi \quad (e^{im\phi} \to (-1)^m e^{im\phi}).
\end{alphaAlign}
Using this and the definition of the spherical harmonics we can show that
\begin{equation}
    Y_l^m = (-1)^l Y_l^m.
\end{equation}
Therefore, we get the transformation under parity to be
\begin{equation}
    \pi \k{\alpha; l, m} = (-1)^l \k{\alpha; l, m}.
\end{equation}

We will now consider the parity properties of the energy eigenkets. A very important theorem reads:
\paragraph{Theorem} Suppose
\begin{equation}
    \comm{H}{\pi} = 0
\end{equation}
and $\k{n}$ is a nondegenerate eigenket of $H$ with eigenvalue $E_n$:
\begin{equation}
    H\k{n} = E_n \k{n};
\end{equation}
then $\k{n}$ is also a parity eigenket.
\paragraph{Proof} First consider the state
\begin{equation}
    \k{\alpha} = \frac{1}{2}(1\pm \pi)\k{n}.
\end{equation}
Then we show
\begin{equation}
    \pi\k{\alpha} = \frac{1}{2}(\pi \pm \pi^2)\k{n} = \frac{1}{2}(\pi \pm 1)\k{n} = \pm \k{\alpha},
\end{equation}
and thus $\alpha$ is a parity eigenket. Since $H$ and $\pi$ commute it is easy to also show that $\k{\alpha}$ is also an energy eigenket with eigenvalue $E_n$. Thus, $\k{n}$ and $\k{\alpha}$ must be the same up to a multiplicative constant, and $\k{n}$ must therefore be a parity eigenket with eigenvalue $\pm 1$. Note also that despite the $\pm$, $\k{\alpha}$ defines just one state. \qed

We now look at the harmonic oscillator as an example. The ground state $\k{0}$, has a Gaussian wave function, which is even under $\vec{x}' \to -\vec{x}'$, and thus $\k{0}$ has even parity. The first excited state
\begin{equation}
    \k{1} = a^\dagger\k{0},
\end{equation}
must have odd parity since $a^\dagger$ is linear in $x$ and $p$, both of which are odd. Continuing this line of reasoning we can show that the $n$th state of the harmonic operator $\k{n}$ has the parity $(-1)^n$. Important to note here is that this result relies on the assumption of non-degeneracy. 

Another example is the momentum eigenket. Since momentum anticommutes with parity, the momentum eigenket is not a parity eigenket, even though the free particle Hamiltonian is invariant under parity. The reason being that we have a degeneracy with $\k{\vec{p}'}$ and $\k{-\vec{p}'}$ having the same energy. Interestingly we can construct eigenkets
\begin{equation}
    \frac{1}{\sqrt{2}}(\k{\vec{p}}' \pm \k{-\vec{p}'})
\end{equation}
which are parity eigenkets. This is in wave functions, equivalent to saying that the plane wave function $e^{i\vec{p}'\cdot\vec{x}'/\hbar}$ does not have definite parity, while the wave functions $\cos(\vec{p}'\cdot\vec{x}'/\hbar)$ and $\sin(\vec{p}'\cdot\vec{x}'/\hbar)$ do.

\subsubsection{Symmetrical Double-Well Potential}
We consider a symmetrical double-well potential. The Hamiltonian is invariant under parity and the two lowest solutions we call the symmetrical state $\k{S}$ and the antisymmetrical state $\k{A}$. These two states are simultaneous eigenkets of both $H$ and $\pi$, and we have
\begin{equation}
    E_A > E_S.
\end{equation}

We form the states
\begin{alphaAlign}
    \k{R} &= \frac{1}{\sqrt{2}} ( \k{S} + \k{A})\\
    \k{L} &= \frac{1}{\sqrt{2}} ( \k{S} - \k{A}).
\end{alphaAlign}
These two states are neither parity states, being interchanged under parity, or energy eigenstates. They are examples of nonstationary states. Consider that at $t=0$ we have the state $\k{R}$, then at a later time we have
\begin{alphaAlign}
    \k{R, t_0=0; t} &= U(t,0) \k{R, 0; 0}\\
    &=\frac{1}{\sqrt{2}} ( e^{-i E_S t/\hbar}\k{S} + e^{i E_A t/\hbar}\k{A})\\
    &= \frac{e^{- i E_S t/\hbar}}{\sqrt{2}} (\k{S} + e^{i(E_A - E_S) t /\hbar}\k{A}.
\end{alphaAlign}
At a time $t = T/2 = \pi\hbar/(E_A-E_S)$ we find the system in $\k{S}$. At the time $t=T$, we have $\k{R}$ again, and we continue oscillating between these two states with angular frequency
\begin{equation}
    \omega = \frac{E_A-E_S}{\hbar}.
\end{equation}
This can be seen as tunnelling between the left and right states. If we increase the middle barrier to infinity, the states $\k{S}$ and $\k{A}$ become degenerate. Thus, $\k{L}$ and $\k{R}$ are now energy eigenkets, but they are not parity eigenkets. The infinitely high barrier makes tunnelling impossible. Interestingly, since we have degeneracy, the ground state can be asymmetrical, even though the Hamiltonian is symmetrical under parity.

\subsubsection{Parity-Selection Rule}
We consider the parity eigenstates $\k{\alpha}$ and $\k{\beta}$, such that
\begin{alphaAlign}
    \pi\k{\alpha} &= \epsilon_\alpha \k{\alpha}\\
    \pi\k{\beta} &= \epsilon_\beta \k{\beta},
\end{alphaAlign}
where $\epsilon_\alpha, \epsilon_\beta$ are the parity eigenvalues $\pm 1$. Unless $\epsilon_\alpha = -\epsilon_\beta$, we have
\begin{equation}
    \bkop{\beta}{\vec{x}}{\alpha} = 0.
\end{equation}
Thus, the odd-parity operator $\vec{x}$ connects parity states of different parity. There is a similar argument due to Wigner
\begin{equation}
    \int d\tau \psi_\beta^* \vec{x}\psi_\alpha = 0,
\end{equation}
if $\psi_b$ and $\psi_a$ have the same parity. This rule was known before the birth of quantum mechanics, from analysis of spectral lines, and called Laporte's rule. Wigner then showed that this rule was due to parity selection. From Wigner's rule it follows that if the Hamiltonian is invariant under parity, then nondegenerate energy eigenstates cannot possess a permanent electric dipole moment, since they are also parity eigenstates.

Generally, operators that have odd parity, have nonvanishing matrix elements only between states of opposite parity, that is, they connect opposite parity states. For operators with even parity, the reverse is true, they connect same parity states.

\subsubsection{Parity Nonconservation}
The Hamiltonian for the weak interaction is not invariant under parity, and as such parity is not conserved through the weak interaction. We can therefore have decays with final states being superpositions of opposite parity states, and the observables can depend on pseudoscalars.

The first experiment that discovered this had a decay rate depending on $\expval{\vec{S}}\cdot \vec{P}$. The decay in question was $\prescript{60}{}{\text{Co}} \to \prescript{60}{}{\text{Ni}} + e^- + \bar{\nu}_e$, where the observables are the $\vec{S}$ for the cobalt nucleus and $\vec{p}$ is the momentum of the emitted electron. 

Since the parity is not conserved in the weak interaction, what was once thought of as pure nuclear and atomic states, are parity mixtures.

\subsection{Lattice Translation as a Discrete Symmetry}
We will consider the discrete symmetry operation that is lattice transformation, with important applications in solid state physics. Consider a periodic potential in one dimension such that $V(x\pm a) = V(x)$. This can be thought of as a chain of positive ions, and an electron moving through this lattice. The Hamiltonian is not generally invariant under a translation $\tau(l)$ with $l$ arbitrary, where $\tau(l)$ has the properties
\begin{alphaAlign}
    \tau^\dagger(l) x \tau(l) = x + l\\
    \tau(l)\k{x'} = \k{x'+l}.
\end{alphaAlign}
However, when $l=a$, with $a$ the lattice spacing we have
\begin{equation}
    \tau^\dagger(a) V(x)\tau(a) = V(x+a) = V(x).
\end{equation}
The kinetic energy part of the Hamiltonian is, of course, invariant under any type of displacement, so we have
\begin{equation}
    \tau^\dagger(a)H\tau(a) = H,
\end{equation}
and also $\tau(a)$ is unitary, with
\begin{equation}
    \comm{H}{\tau(a)} = 0,
\end{equation}
such that both $H$ and $\tau(a)$ can be diagonalized simultaneously. Note that $\tau(a)$ is not Hermitian, and we expect the eigenvalue to be complex modulus 1.

We will start by considering a periodic potential with the potential between two sites going to infinity. A state completely localized in a lattice site is a good ground state candidate. Assuming that the particle is localized at the $n$th site, denoted by $\k{n}$ with the ground state energy $E_0$ such that $H\k{n} = E_0\k{n} $. The wave function $\bk{x'}{n}$ is finite only on the $n$th site, but we note that similar states exists on any state $n$ where $n\in(-\infty,\infty)$.

The ket $\k{n}$ is obviously not an eigenket to $\tau(a)$ since
\begin{equation}
    \tau(a)\k{n} = \k{n+1}.
\end{equation}
We see once again that even though $\tau(a)$ and $H$ commute we do not have a simultaneous energy eigenket, consistent with us having an infinite-fold degeneracy. We want to find some ket that is a simultaneous eigenket of both $H$ and $\tau(a)$.

A similar procedure as performed earlier with the states $\k{L}$ and $\k{R}$ will be done here, where even though they were not parity eigenkets, some linear combination were. The linear combination
\begin{equation}
    \k{\theta} = \sum_{n=-\infty}^{\infty} e^{in\theta}\k{n},
\end{equation}
with $\theta\in(-\pi,\pi]$ is formed. This is obviously an eigenket of $H$, since $\k{n}$ is, and applying the lattice translation operator
\begin{equation}
    \tau(a)\k{\theta} = \sum_{-\infty}^{\infty} e^{in\theta}\k{n+1} = \sum_{n=-\infty}^{\infty} e^{i(n-1)\theta} \k{n} = e^{-i\theta}\k{\theta}.
\end{equation}
Thus, $\k{\theta}$ is a simultaneous eigenket of $H$ and $\tau(a)$.

Returning to the lattice potential with finite height, albeit still tall enough. It is still possible to construct the kets $\k{n}$ with property $\tau(a)\k{n} = \k{n+1}$, but the wave function $\bk{x'}{n}$, can now leak into neighbouring sites due to quantum tunnelling effects. Assuming that the potential is high enough, we can ignore the leakage into sites other than the nearest neighbours. That is, we get the Hamiltonian
\begin{equation}
    \bkop{n}{H}{n} = E_0 \quad \text{and} \quad \bkop{n'}{H}{n} = -\Delta \quad \text{only if}\quad n' = \pm 1.
\end{equation}
This is called the tight-binding approximation in solid state physics. We thus get
\begin{equation}
    H\k{n} = E_0\k{n} - \Delta \k{n+1} - \Delta \k{n-1}.
\end{equation}
Note that $\k{n}$ is no longer an energy eigenket.

We form that same linear combination as before
\begin{equation}
    \k{\theta} = \sum_{n=-\infty}^{\infty} e^{in\theta}\k{n}.
\end{equation}
This is still an eigenket to $\tau(a)$ since the steps performed earlier still holds. However, since $\k{n}$ is not an energy eigenket any more, it is not obvious if $\k{\theta}$ is. We test this by applying $H$
\begin{alphaAlign}
    H\k{\theta} &= H \sum_n e^{in\theta} \k{n}\\
    &= \sum_n e^{in\theta} H\k{n}\\
    &= \sum_n e^{in\theta} (E_0\k{n} - \Delta\k{n+1} - \Delta\k{n-1})\\
    &= E_0\sum_n e^{in\theta}\k{n} -\Delta \sum_n e^{in\theta} \ab(e^{-i\theta} + e^{i\theta})\k{n}\\
    &= (E_0 -2\Delta\cos\theta)\sum_n e^{in\theta} \k{n}\\
    &= (E_0 - 2\Delta\cos\theta) \k{\theta}.
\end{alphaAlign}
What is interesting here is that the eigenvalue depends on the parameter $\theta$ which it did not do for the infinite potential case. As $\Delta$ becomes non-zero the degeneracy is removed, and we get a continuous spectrum of energies in the range $E\in[E_0 - 2\Delta, E_0 + 2\Delta]$.

Now, we want to apply some physical significance to the parameter $\theta$, and we will therefore study the wave function $\bk{x'}{\theta}$. The wave function for the lattice translated state $\tau(a)\k{\theta}$ is 
\begin{equation}
    \bkop{x'}{\tau(a)}{\theta} = \bk{x'-a}{\theta}
\end{equation}
or
\begin{equation}
    \bkop{x'}{\tau(a)}{\theta} = e^{-i\theta}\bk{x'}{\theta},
\end{equation}
depending on if we let the operator act on the bra or ket. Thus,
\begin{equation}
    \bk{x'-a}{\theta} = \bk{x'}{\theta} e^{-i\theta},
\end{equation}
which can be solved by setting
\begin{equation}
    \bk{x'}{\theta} = e^{ikx'}u_k(x'),
\end{equation}
with $\theta=ka$. The function $u_k(x')$ is a periodic function with period $a$. With this we arrive at Bloch's theorem: The wave function of $\k{\theta}$, which is an eigenket of $\tau(a)$, can be written as a plane wave $e^{ikx'}$ times a periodic function with period $a$. This theorem holds even if the tight-binding approximation breaks down.

We can now interpret the meaning of $\k{\theta}$ and its energy eigenvalue. We know that the wave function is a plane wave characterized by the wave vector $k$ and modulated by the periodic function $u_k(x')$. As $\theta = ka$ varies between $-\pi$ and $\pi$ we have that the wave vector $k$ varies between $-\pi/a$ and $\pi/a$. The energy eigenvalue has a $k$ dependence
\begin{equation}
    E(k) = E_0 - 2\Delta \cos ka.
\end{equation}
This is independent of the shape of the potential as long as the tight-binding approximation holds. Note that we have a cutoff for the wave vector at $\abs{k} = \pi/a$. The dispersion curve defined by $E(k)$ forms a continuous energy band between $E_0 - 2\Delta$ and $E_0 + 2\Delta$ called the Brillouin zone. Note also that we have lifted the infinite-fold degeneracy completely.

When considering multiple electrons, following the Pauli principle, as opposed to just a single electron as considered here, we can find many of the features of metals, semiconductors and insulators.

\subsection{The Time-Reversal Discrete Symmetry}
We will now explore the discrete symmetry operator that is time reversal. The term time reversal is a misnomer, what we mean are more accurately described by the term reversal of motion. This is the term used by Wigner when he formulated time reversal.

We start from a classical point of view. Imagine the trajectory of a particle in a certain force field. Suppose now that at $t=0$ the particle stop and reverse its motion, that is $\vec{p}|_{t=0} \to -\vec{p}|_[t=0]$. The particle thus moves backwards along the same path. We can say, more formally, that if $\vec{x}(t)$ is a solution to
\begin{equation}
    m\ddot{\vec{x}}(t) = -\nabla V(\vec{x}),
\end{equation}
then $\vec{x}(-t)$ is also a valid solution. Note that this is not true for non-conserving forces, or dissipative forces.

Can we then tell if we are either following a path A in reverse, or a path B in forward motion? If the field in question is a magnetic field it might be possible. Imagine an electron spiralling in a magnetic field, it might be possible, from a macroscopic view, to get a sense of rotation by noting the labelling of the magnetic poles. However, if we also reverse the current that generates $\vec{B}$, then the situations would be symmetrical. That is, we can say formally that Maxwells equations, e.g.
\begin{equation}
    \nabla \cdot \vec{E}, \quad \nabla\times\vec{B} - \frac{1}{c} \partial_t \vec{E}=\frac{4\pi\vec{j}}{c}, \quad \nabla\times\vec{E} = -\frac{1}{c}\partial_t\vec{B},
\end{equation}
and the Lorentz force 
\begin{equation}
    \vec{F} = e \ab[\vec{E} + \frac{1}{c}(\vec{v}\times\vec{B})]
\end{equation}
are invariant under the transformation $t\to -t$, if we also have 
\begin{equation}
    \vec{E}\to \vec{E}, \quad \vec{B}\to -\vec{B}, \quad \rho \to \rho, \quad \vec{j} \to -\vec{j}, \quad\text{and}\quad \vec{v}\to -\vec{v}.
\end{equation}

Moving away from the classical point of view, and instead looking to wave mechanics, the Schrödinger equation reads
\begin{equation}
    i\hbar \partial_t \psi = \ab(-\frac{\hbar^2}{2m}\nabla^2 + V)\psi.
\end{equation}
If we suppose that $\psi(\vec{x}, t)$ is a solution, then it is easy to see that $\psi(\vec{x}, -t)$ is not a solution due to the first time derivative. We do see however, that $\psi^*(\vec{x}, -t)$ is a solution. For an energy eigenstate
\begin{equation}
    \psi(\vec{x}, t) = u_n(\vec{x}) e^{-iE_n t/\hbar}, \quad\text{and}\quad \psi^*(\vec{x}, -t) = u_n^*(\vec{x}) e^{-iE_n t/\hbar}.
\end{equation}
We conjecture that time reversal and complex conjugation are related. 

